\LARGE Performance Evaluation of Xenomai 3

### 2015-10-20 paper.tex
Not include eps files anymore. Increase latency graph legends spacing and fonts.
author: adjust the order to be consistent
make \cite work again
Modify experiment paragraphs

\author[*]{\large
{\bf Ching-Chun (Jim) Huang}\thanks{jserv@ccns.ncku.edu.tw}}
\author[*]{\large
{\bf Che-Kang Wu}\thanks{an4006048$@$mail.ncku.edu.tw}}
\author[**]{\large
{\bf Chan-Hsiang Lin}\thanks{r04943031$@$ntu.edu.tw}}
\affil[*]{Department of Computer Science and Information Engineering,
\newline
National Cheng Kung University, Taiwan
\newline
No.1, University Road, Tainan City 701, Taiwan (R.O.C.)}
\affil[**]{Department of Electrical and Electronic Engineering,
\newline
National Taiwan University
\newline
No.1, Sec. 4, Roosevelt Road, Taipei, Taiwan (R.O.C.)}
\date{}

@@

\section{Experiment}

\subsection{Purpose}
1. Compare the performance of Xenomai with baseline Linux kernel and PREEMPT\_RT patched kernel
2. Compare the performance of Xenomai-2 and Xenomai-3
\subsection{Environment Setup}

\subsubsection{Hardware}
For the testing platform, we used the BeagleBone Black. BeagleBone Black is an effective embedded platform with ARM Cortex-A8 processor, capable of running Linux and Xenomai (both 2 and 3), consisting of a single processor, a 512 MB memory, 32kB L1 caches and a 256kB L2 cache.

\subsubsection{Linux Kernel}
We installed a Debian Linux image\cite{debian} on the beaglebone, with hardware-specific linux kernel from\cite{kernel}. Kernel version 3.14.39 were chosen due to it's compatibility with all kernel configurations.
Four distinct kernel configurations are used for the experiment, as summarized in table\ref{table:config}. Kernel are patched using git merge first, and remaining conflicts are resolved manually.

@@

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & source\\ \hline
mainline   & none & n/a \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (2015-07-13 HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0  & \cite{p-ipipe} \cite{git-x3} \\ \hline
\end{tabular}

@@

  The gravity values is the shortest time the platform needs to deliver an IRQ to a Xenomai interrupt handler,  a RTDM task running in kernel space, or a user-space task. It is used to differentiate timers on the target context they activate, among IRQ(handler), kernel and user threads, anticipating the next timer shot accordingly, so that such context is activated as close as possible to the ideal time. The gravity values can be set in the file \begin{verbatim}/proc/xenomai/latency on Xenomai-2, /proc/xenomai/clock/coreclk
\end{verbatim} on Xenomai-3 with root privilege. Furthermore, Xenomai-3 has an autotune function that runs a series of internal calibration tests for estimating the most appropriate gravity values for its real-time clock timer, retaining the final values.
Because Xenomai-2 doesn't have an autotune function and it has only one gravity value for all the 
coreclocks, we decided to manually tune all the coreclock values in both Xenomai-2 and Xenomai-3 to zero, so that we can later compare the results on the same basis.

@@

\subsubsection{Scenario}
The goal of the experiment is to evaluate the realtime performances of Mainline Linux, PREEMPT\_RT, 
Xenomai-2 and Xenomai-3. Also, we want to find out whether the newly released Xenomai-3.0 has improved in realtime capability, compared to Xenomai-2.6.4, the latest stable version of Xenomai-2.
\subsubsection{Tools}
Cyclictest is a commonly-used latency benchmark. It is compatible with all the kernel setups we use (the recently Xenomai-3.0 release add cyclictest to its testsuite), therefore it's suitable for the purpose of measuring latencies for all setups.
Latency is a test program for the Xenomai test suite. It can measure latencies under 3 modes - user task, kernel task, or timer IRQ. It launches the test procedure in one of the user-specified contexts. In the user mode, the measurement task is handled by the 'latency' function in the program, which periodically waits and fetches the current time when resuming. Latency is then measured by calculating the difference of the actual time elapsed with the predefined period.
\subsubsection{Test Procedure}
Each experiment was run on the same hardware. Each was run for 20 minutes to generate approximately 1,000,000 samples. Test scripts were written to facilitate the experiment process.
\subsubsection{Test Load}
As previously mentioned, we measure latencies under IDLE and CPU-STRESSED conditions. In the CPU-STRESSED condition, we called the "stress" process to exert pressure on the CPU.
\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=8cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=8cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-tress}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=8cm]{img/latency_idle.jpg}
\caption{Latency with no stress.}
\label{latency-idle}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=8cm]{img/latency_stress.jpg}
\caption{Latency with no stress.}
\label{latency-stress}
\end{center}
\end{figure}

\epsin{cyclic-idle}{80}{fig2:f2}{Cyclictest Idle}
\epsin{cyclic-cpu}{80}{fig3:f3}{Cyclictest stressed}

\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
\end{tabular}

@@

\bibliographystyle{plain}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
\end{thebibliography}

>>>>

\author[*]{\large
{\bf Ching-Chun (Jim) Huang}\thanks{jserv@ccns.ncku.edu.tw}}
\author[**]{\large
{\bf Chan-Hsiang Lin}\thanks{r04943031$@$ntu.edu.tw}}
\author[*]{\large
{\bf Che-Kang Wu}\thanks{an4006048$@$mail.ncku.edu.tw}}
\affil[*]{Department of Computer Science and Information Engineering,
\newline
National Cheng Kung University, Taiwan
\newline
No.1, University Road, Tainan City 701, Taiwan (R.O.C.)}
\affil[**]{Department of Electrical and Electronic Engineering,
\newline
National Taiwan University
\newline
No.1, Sec. 4, Roosevelt Road, Taipei, Taiwan (R.O.C.)}
\date{}

@@

\section{Experiment}
There are two objectives we wish to complete in our experiment. First, compare the performance of Xenomai with baseline Linux kernel and PREEMPT\_RT patched kernel to decide which configuration has better realtime capability. Second, try to validate whether Xenomai-3 has improved its realtime performance. The details are described in following sub-chapters. 
\subsection{Environment Setup}
Because building and transporting operating systems on embedded devices is complicated. In order for other people to re-do the experiment, it needs proper setup and documentation. We will elaborate on different parts of the experiment setup. 
\subsubsection{Hardware}
For the testing platform, we used the BeagleBone Black. BeagleBone Black is an effective embedded platform with ARM Cortex-A8 processor, capable of running Linux and Xenomai (both 2 and 3), consisting of a single processor, a 512 MB memory, 32kB L1 caches and a 256kB L2 cache.  
\subsubsection{Linux Kernel}
We installed a Debian Linux image\cite{debian} on the beaglebone, with hardware-specific linux kernel from\cite{kernel}. Kernel version 3.14.39 were chosen due to it's compatibility with all kernel configurations.
Four distinct kernel configurations are used for the experiment, as summarized in table\ref{table:config}. Kernel are patched using git merge first, and remaining conflicts are resolved manually.

@@

\begin{tabular}{|l| p{3cm} |l|}
\hline
Kernel & Patches & source\\ \hline
mainline   & none & n/a \\ \hline
PREEMPT\_RT & patch-3.14.39-rt37 & \cite{p-rt} \\ \hline
Xenomai 2  & ipipe-core-3.14.39-arm-9 + xenomai v2.6.4 (2015-07-13 HEAD) & \cite{p-ipipe} \cite{git-x2.6} \\ \hline
Xenomai 3  & ipipe-core-3.14.39-arm-9 + xenomai v3.0  & \cite{p-ipipe} \cite{git-x3} \\ \hline
\label{table:config}
\end{tabular}

@@

  The gravity values is the shortest time the platform needs to deliver an IRQ to a Xenomai interrupt handler,  a RTDM task running in kernel space, or a user-space task. It is used to differentiate timers on the target context they activate, among IRQ(handler), kernel and user threads, anticipating the next timer shot accordingly, so that such context is activated as close as possible to the ideal time. The gravity values can be set in the file \begin{verbatim}/proc/xenomai/latency\end{verbatim} on Xenomai-2, and in the file \begin{verbatim}/proc/xenomai/clock/coreclk
\end{verbatim} on Xenomai-3 with root privilege. Furthermore, Xenomai-3 has an autotune function that runs a series of internal calibration tests for estimating the most appropriate gravity values for its real-time clock timer, retaining the final values.
Because Xenomai-2 doesn't have an autotune function and it has only one gravity value for all the 
coreclocks, we decided to manually tune all the coreclock values in both Xenomai-2 and Xenomai-3 to zero, so that we can later compare the results on the same basis.

@@

\subsubsection{Scenario}
The goal of the experiment is to evaluate the realtime performances of Mainline Linux, PREEMPT\_RT, 
Xenomai-2 and Xenomai-3. Also, we want to find out whether the newly released Xenomai-3.0 has improved in realtime capability, compared to Xenomai-2.6.4, the latest stable version of Xenomai-2.
\subsubsection{Tools}
Cyclictest is a commonly-used latency benchmark. It is compatible with all the kernel setups we use (the recently Xenomai-3.0 release add cyclictest to its testsuite), therefore it's suitable for the purpose of measuring latencies for all setups.
Latency is a test program for the Xenomai test suite. It can measure latencies under 3 modes - user task, kernel task, or timer IRQ. It launches the test procedure in one of the user-specified contexts. In the user mode, the measurement task is handled by the 'latency' function in the program, which periodically waits and fetches the current time when resuming. Latency is then measured by calculating the difference of the actual time elapsed with the predefined period.
\subsubsection{Test Procedure}
Each experiment was run on the same hardware. Each was run for 20 minutes to generate approximately 1,000,000 samples. Test scripts were written to facilitate the experiment process.
\subsubsection{Test Load}
As previously mentioned, we measure latencies under IDLE and CPU-STRESSED conditions. In the CPU-STRESSED condition, we called the "stress" process to exert pressure on the CPU.
\subsection{Experiment Result}
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_idle.jpg}
\caption{Cyclictest with no stress.}
\label{cyclictest-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/cyclictest_stress.jpg}
\caption{Cyclictest with stress.}
\label{cyclictest-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-idle.png}
\caption{Latency with no stress on Xenomai-2.}
\label{x2-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x2-cpu.png}
\caption{Latency with stress on Xenomai-2.}
\label{x2-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-idle.png}
\caption{Latency with no stress on Xenomai-3.}
\label{x3-latency-idle}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\begin{center}
\includegraphics[width=6cm]{img/x3-cpu.png}
\caption{Latency with stress on Xenomai-3.}
\label{x3-latency-stress}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tabular}{|l|c|r|r|}
\hline
\multirow{2}{*}{} & \multicolumn{3}{c|}{user} \\ \cline{2-4} 
 & min & \multicolumn{1}{c|}{avg} & \multicolumn{1}{c|}{max} \\ \hline
Xenomai 2 & 8.541 & 9.458 & 34.583 \\ \hline
Xenomai 3 & 8.043 & 8.853 & 33.023 \\ \hline
& \multicolumn{3}{c|}{kernel}      \\ \hline
Xenomai 2 & 6.965 & 7.708 & 27.821 \\ \hline
Xenomai 3 & 5.567 & 6.479 & 25.577 \\ \hline
& \multicolumn{3}{c|}{timer-irq}   \\ \hline
Xenomai 2 & 2.129 & 2.654 & 11.343 \\ \hline
Xenomai 3 & 2.586 & 3.079 & 10.031 \\ \hline
\end{tabular}

@@

\bibliographystyle{unsrt}
\begin{thebibliography}{9}%use this if you have <=9 bib refs
	\bibitem {rtai}{\it https://www.rtai.org/}
	\bibitem {sel4}{\it http://ssrg.nicta.com.au/projects/seL4/}
	\bibitem {debian}{\it http://debian.beagleboard.org/images/bone-debian-7.5-2014-05-14-2gb.img.xz}
	\bibitem {kernel}{\it https://github.com/beagleboard/linux}
	\bibitem {p-rt}{\it https://www.kernel.org/pub/linux/kernel/projects/rt/3.14/}
	\bibitem {p-ipipe}{\it http://download.gna.org/adeos/patches/v3.x/arm/ipipe-core-3.14.39-arm-9.patch}
	\bibitem {git-x2.6}{\it http://git.xenomai.org/xenomai-2.6.git/}
	\bibitem {git-x3}{\it http://git.xenomai.org/xenomai-3.git/}
\end{thebibliography}
